---
title: "happiness analysis"
author: "Damiano Taricone - Alessandro Carmellini - Francesco Poggi"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

## Library Imports

```{r}
library(mclust)
library(magrittr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(NbClust)
library(cluster)
library(dplyr)
library(factoextra)
library(lme4) 
library(haven) 
library(clustvarsel)
library(plotly)
library(corrplot)
library(reshape2)
library(plotly)
```

## Descriptive Analysis

Dataset Loading

```{r}
data <- read.csv("2018.csv")
```

Correlation Plot

```{r}
plot(data[,-c(1,2)]) 
```

Missing Values

```{r}
missing_values_per_variable <- colSums(is.na(data))
print(missing_values_per_variable)
```

Barplot of Scores

```{r}
Scoreplot <- ggplot(data, aes(x = reorder(`Country.or.region`, -Score), y = Score, fill = Score)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  labs(title = "Country or Region and Happiness Score",
       x = "Country or region",
       y = "Happiness Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_gradient(low = "darkblue", high = "lightcoral") 

print(Scoreplot)
```

Top and Bottom 20 Countries

```{r}
top_countries <- head(data[order(-data$Score), ], 20)
print(top_countries)

# Bottom 20 least happy countries
bottom_countries <- head(data[order(data$Score), ], 20)
print(bottom_countries)

top_20 <- head(data[order(-data$Score), ], 20)
bottom_20 <- head(data[order(data$Score), ], 20)
```

Variable Composition

```{r}
data$Sum_Variables <- rowSums(data[, c("GDP.per.capita", "Social.support", "Healthy.life.expectancy", "Freedom.to.make.life.choices", "Generosity", "Perceptions.of.corruption")], na.rm = TRUE)
melted_data <- reshape2::melt(data, id.vars = c("Country.or.region", "Sum_Variables"), measure.vars = c("GDP.per.capita", "Social.support", "Healthy.life.expectancy", "Freedom.to.make.life.choices", "Generosity", "Perceptions.of.corruption"))
my_palette <- c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854", "#ffd92f")

# Grafico a barre con colori diversi per ogni variabile
Compplot <- ggplot(melted_data, aes(x = reorder(`Country.or.region`, -Sum_Variables), y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = my_palette) +
  labs(title = "Grafico composizione Score",
       x = "Country",
       y = "Valore") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # Etichette a 90 gradi
        legend.position = "top", legend.title = element_blank())

print(Compplot)
```

# Regression

Regression Analysis GDP and Score Relationship

```{r}
scatter_plot <- ggplot(data, aes(x = `GDP.per.capita`, y = Score, text = `Country.or.region`)) +
  geom_point(aes(color = ifelse(`Country.or.region` %in% top_20$`Country.or.region`, "Top 20 Score Alto", 
                                ifelse(`Country.or.region` %in% bottom_20$`Country.or.region`, "Top 20 Score Basso", "Altri Paesi"))),
             size = 2, alpha = 0.8) +
  scale_color_manual(values = c("gray", "red", "blue"), 
                     labels = c("Altri Paesi", "Top 20 Score Alto", "Top 20 Score Basso")) +
  labs(title = "Rapporto tra Score e GDP",
       x = "GDP per capita",
       y = "Score") +
  theme_minimal() +
  guides(color = FALSE)  

plotly11 <- ggplotly(scatter_plot, tooltip = "text")
print(plotly11)
```

Score and Corruption Perception

```{r}
Scatter_corr <- ggplot(data, aes(x = `Perceptions.of.corruption`, y = Score, text = `Country.or.region`)) +
  geom_point(aes(color = ifelse(`Country.or.region` %in% top_20$`Country.or.region`, "Top 20 Score Alto", 
                                ifelse(`Country.or.region` %in% bottom_20$`Country.or.region`, "Top 20 Score Basso", "Altri Paesi"))),
             size = 2, alpha = 0.8) +
  scale_color_manual(values = c("gray", "red", "blue"), 
                     labels = c("Altri Paesi", "Top 20 Score Alto", "Top 20 Score Basso")) +
  labs(title = "Rapporto tra Score e Perceptions of corruption",
       x = "Perceptions of corruption",
       y = "Score") +
  theme_minimal() +
  guides(color = guide_legend(title = NULL))

plotly22 <- ggplotly(Scatter_corr, tooltip = "text")
print(plotly22)
```

Score and Social Support

```{r}
Scatter_social <- ggplot(data, aes(x = `Social.support`, y = Score, text = `Country.or.region`)) +
  geom_point(aes(color = ifelse(`Country.or.region` %in% top_20$`Country.or.region`, "Top 20 Score Alto", 
                                ifelse(`Country.or.region` %in% bottom_20$`Country.or.region`, "Top 20 Score Basso", "Altri Paesi"))),
             size = 2, alpha = 0.8) +
  scale_color_manual(values = c("gray", "red", "blue"), 
                     labels = c("Altri Paesi", "Top 20 Score Alto", "Top 20 Score Basso")) +
  labs(title = "Rapporto tra Score e Social support ",
       x = "Social Support",
       y = "Score") +
  theme_minimal() +
  guides(color = guide_legend(title = NULL))

plotly33 <- ggplotly(Scatter_social, tooltip = "text")
print(plotly33)
```

Freedom and GDP

```{r}
Scatter_free_GDP <- ggplot(data, aes(x = `Freedom.to.make.life.choices`, y = `GDP.per.capita`, text = `Country.or.region`)) +
  geom_point(aes(color = ifelse(`Country.or.region` %in% top_20$`Country.or.region`, "Top 20 Score Alto", 
                                ifelse(`Country.or.region` %in% bottom_20$`Country.or.region`, "Top 20 Score Basso", "Altri Paesi"))),
             size = 2, alpha = 0.8) +
  scale_color_manual(values = c("gray", "red", "blue"), 
                     labels = c("Altri Paesi", "Top 20 Score Alto", "Top 20 Score Basso")) +
  labs(title = "Rapporto tra Freedom e GDP ",
       x = "Freedom",
       y = "GDP") +
  theme_minimal() +
  guides(color = guide_legend(title = NULL))

plotly44 <- ggplotly(Scatter_free_GDP, tooltip = "text")
print(plotly44)
```

Freedom and Generosity

```{r}
Scatter_free_Gen <- ggplot(data, aes(x = `Freedom.to.make.life.choices`, y = `Generosity`, text = `Country.or.region`)) +
  geom_point(aes(color = ifelse(`Country.or.region` %in% top_20$`Country.or.region`, "Top 20 Score Alto", 
                                ifelse(`Country.or.region` %in% bottom_20$`Country.or.region`, "Top 20 Score Basso", "Altri Paesi"))),
             size = 2, alpha = 0.8) +
  scale_color_manual(values = c("gray", "red", "blue"), 
                     labels = c("Altri Paesi", "Top 20 Score Alto", "Top 20 Score Basso")) +
  labs(title = "Rapporto tra Freedom e Generosity ",
       x = "Freedom",
       y = "Generosity") +
  theme_minimal() +
  guides(color = guide_legend(title = NULL))  # Rimuovere il titolo della legenda


plotly55 <- ggplotly(Scatter_free_Gen, tooltip = "text")
print(plotly55)
```

## Boxplots and Standardization

Original Boxplot

```{r}
BOXPLOTORIG <- data[,-c(1,2)] %>%
  gather(Attributes, values) %>%
  ggplot(aes(x = reorder(Attributes, values, FUN = median), y = values, fill = Attributes)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title = "Boxplot of All Variables") +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  coord_flip()

print(BOXPLOTORIG)
```

Standardized Boxplot

```{r}
datastan <- as.data.frame(scale(as.data.frame(data[,-c(1,2)])))

# Boxplot of standardized variables
BOXSTAND <- datastan %>%
  gather(Attributes, values) %>%
  ggplot(aes(x = reorder(Attributes, values, FUN = median), y = values, fill = Attributes)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title = "Boxplot of Standardized Variables") +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.title.x = element_blank()) +
  coord_flip()

print(BOXSTAND)
```

## Clustering Analysis

Hierarchical Clustering

```{r}
# Hierarchical clustering method 
DistEuc = dist(datastan, method = "euclidean") # Using a dataset without categorical variables

EucWard = hclust(DistEuc, method = "ward.D2", members = NULL) # Choosing Ward method

# Plotting the dendrogram
plot(EucWard)
rect.hclust(EucWard, k = 3, border = "red")
fviz_dend(EucWard, k = 3, show_labels = F, rect = T) # Colored dendrogram
TaglioEucWard = cutree(EucWard, k = 3)

# Plotting the data with cluster colors
plot(datastan, col= TaglioEucWard)#il gruppo piu piccolo si distacca sempre, potrebbero essere anche solo 2 gruppi, ma con 3 dovremmo 
```

### K-means Clustering

```{r}
# K-means clustering
fviz_nbclust(datastan, kmeans, method = "silhouette") # Proposing 3 clusters based on silhouette method
fviz_nbclust(datastan, kmeans, method = "wss") # Proposing 3 clusters based on within-sum-of-squares

kmeans_result <- kmeans(datastan, 3, nstart = 100)

# Visualization of k-means clusters
cluster_visualization <- fviz_cluster(kmeans_result, data = datastan)
print(cluster_visualization)

# Tabulation of cluster occurrences
cluster_occurrences <- table(kmeans_result$cluster)
print(cluster_occurrences)

# Creating an additional column containing cluster membership
datastan$cluster <- kmeans_result$cluster
write.csv(datastan, "data_with_cluster.csv", row.names = FALSE)
```

Distribution of Score

```{r}
# Histogram of Score variable
x <- data$Score
hist(x, breaks = 20, freq = FALSE)

# Clustering using Mclust
CLUSTERING <- Mclust(x, G = 3, modelNames = "V")
SEQ <- seq(min(x), max(x), length = 1500)

DENS1 <- dnorm(SEQ, mean = CLUSTERING$parameters$mean[1], 
               sd = sqrt(CLUSTERING$parameters$variance$sigmasq[1]))
DENS2 <- dnorm(SEQ, mean = CLUSTERING$parameters$mean[2], 
               sd = sqrt(CLUSTERING$parameters$variance$sigmasq[2]))
DENS3 <- dnorm(SEQ, mean = CLUSTERING$parameters$mean[3], 
               sd = sqrt(CLUSTERING$parameters$variance$sigmasq[3]))

P1 <- CLUSTERING$parameters$pro[1]
P2 <- CLUSTERING$parameters$pro[2]
P3 <- CLUSTERING$parameters$pro[3]

DENSTOT <- P1 * DENS1 + P2 * DENS2 + P3 * DENS3

# Plotting the density lines
lines(SEQ, DENSTOT)
lines(SEQ, P1 * DENS1, col = 2)
lines(SEQ, P2 * DENS2, col = 3)
lines(SEQ, P3 * DENS3, col = 4)
```

Variable Selection Heatmap of Correlations

```{r}
# Heatmap for variable correlations
print(names(data))
# Selecting only the variables of interest
variables <- c("Score", "GDP.per.capita", "Social.support", "Healthy.life.expectancy", "Freedom.to.make.life.choices", "Generosity", "Perceptions.of.corruption")
correlation_data <- data[, variables]

# Calculating the correlation matrix
cor_matrix <- cor(correlation_data, use = "complete.obs")

# Transforming the correlation matrix into a long format
melted_cor_matrix <- melt(cor_matrix)

# Creating a heatmap with annotations of values
heatmap_correlation <- ggplot(melted_cor_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), vjust = 1) +  # Adding value annotations
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Heatmap of Correlations",
       x = "Variable",
       y = "Variable") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1))

print(heatmap_correlation)
```

Selection with Boruta

## Clusvarsel

```{r}
# Installing and loading Boruta for variable selection
library(Boruta)
library(ranger)

# Boruta analysis for feature selection
boruta_result <- Boruta(Score ~ ., data = datastan[,-c(8,9)], doTrace = 2)
print(boruta_result)

# Plotting the importance of variables
plot(boruta_result, cexAxis = 0.7, las = 2, xlab = "", main = "Boruta Analysis: Importance of Variables")
```

```{r}
datastanclean <- datastan[, -c(8, 9)]

cluststanDBACK <- clustvarsel(datastanclean,G=1:3, direction = "backward") #Seleziona: Perceptions of corruption, Freedom to make life choices, Generosity, Healthy life expectancy
#seleziona G= 3 e variabili 1,3,4,5,7
plot(cluststanDBACK$model)
cluststanDBACK$model$classification
cluststanDBACK$subset

```
